---
title: "Nearest Neighbor Classifier"
author: "Dmytro Fishman"
date: "29 July 2017"
output: html_document
---

# Nearest Neighbor Classifier

Load and subsample the MNIST dataset from the previous lesson

```{r}
library(mnist)
data(mnist)
images <- mnist$train
set.seed(1111) # you can use set.seed() in order to reproduce stochastic results
new_indx <- sample(c(1:nrow(images$x)), size = 2000, replace = FALSE)
labels <- images$y[new_indx]
images <- images$x[new_indx,]
```

Now we will define a distance function...

```{r}
dist <- function(img1, img2) {
  "... your code here ..."
}
```

You can play with the defined function below. Check that the distance between images of the same digit is typically smaller than the distance between images of different digits.

```{r}
print(paste("Distance between images of class", labels[1], "and", labels[10], "is", dist(images[1,], images[10,])))
print(paste("Distance between images of class",labels[1], "and", labels[4], "is", dist(images[1,], images[4,])))
```

Now the actual algorithm. It consists of three steps.
- Compute distances to all points in the dataset
- Find the closest point, and
- Report the corresponding label.

*Exercise*: Implement the three steps.

We shall do it gently. First we pick a sample image that we will be classifying. Let us say this will be an unknown image sent by our friend:

```{r}
unknown_img <- mnist$test$x[1,]
true_label <- mnist$test$y[1]
```

Compute all distances from the `unknown_img` to the images in the dataset

```{r}
# we shall use function apply() which is a way of doing list comprehension in R 
# here we use apply to iterate through images by rows (argument '1') and 
# for each image compute the distance to unknown image
all_distances <- "... your code here ..."
head(all_distances)
```

Now let's find out which image is closest to our `unknown_img`

```{r}
(closest_index <- which.min(all_distances))
```

Almost done, now report a label with index i in labels

```{r}
(predicted_label <- labels[closest_index])
```

Compare it to the true label of the first image in the test labels

```{r}
(predicted_label == true_label)
print(paste("Predicted class for the first image is", predicted_label ,"and the true label is",  true_label))
```

*Exercise!* Use code from the previous practice session to plot the first example and visually confirm it's label

```{r, echo=FALSE}
# first we need to define colors:
colors<-c('black','white')
cus_col<-colorRampPalette(colors=colors)

# Play around with an index of image that you want to visualise
index <- 1
img <- array(mnist$test$x[index,],dim=c(28,28))
img <- img[,28:1]
image(img, col=cus_col(256))
```

Now let's make a function out of the code we have already written, copy paste appropriate lines below

```{r}
classify <- function(unknown_img) {
    all_distances <- apply(images, 1, function(img) dist(unknown_img, img))
    return("... your code here ...")
}
```

Test it to verify that it works

```{r}
print(paste("Predicted class for the first image is", classify(unknown_img),"and the true label is",  true_label))
```

One very popular variation of Nearest Neighbour is K-nearest neighbour. In this algorithm a label for a new instance is chosen by majority vote by k of its nearest neighbors.

The actual algorithm is not very different from vanila nearest neighbour:
- Compute distances to all points in the dataset
- Find the k closest points
- Report the most popular label from these k.

*Exercise*: Implement KNN algorithm reusing some code from dist() and classify() functions.

```{r}
classify_knn <- function(unknown_img, k = 5) {
    # This step we already know from the previous exercises
    all_distances <- apply(images, 1, function(img) dist(unknown_img, img))
    
    # We need to get indexes of K smallest distances 
    # (hint: use functions order() and head())
    knn = "... your code here ..."
    
    # you can print potential predictions
    print(labels[knn])

    # Very small step is left, return the most frequently predicted label
    # (hint: use functions table() and sort())
    return("... your code here ...")
}
```

Test this version of KNN, experiment with different `K`s

```{r}
print(paste("Predicted class for the first image is", classify_knn(unknown_img, k = 4),"and the true label is",  true_label))
```


